Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2021-01-27 spectral radius of matrix new method/thesis/final_article0127.tex
publishing 0127

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
 \usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman
\usepackage{graphicx}
\usepackage{comment}

@@

@@

The spectral radius is an important indicator to specify the relation
 of connected vertices in a graph, so it is meaningful to find a simple
 method to estimate the spectral radius. A simple and excellent
 executable method to estimate the spectral radius has some features,
 first, the bios is minimized, and second, there must be a way to prove
 it sensible. Enumerate these factors and prove it correctly would make this method reliable.

In \cite{chang}, Cheng and Weng give many bounds of the spectral radius of a nonnegative square matrix. And based on their theory and Perron-Frobenius theorem, we give another approach to obtain an upper bound of the spectral radius and apply it on the adjacency matrix of a directed graph.
All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented.\cite{src_files}

@@

 A well-known application of Theorem~\ref{thm:Perron_Frobenius}
 show that if matrix  $C'$ {\it majors} $C$ (in notation $C\leq C'$),
 i.e. $c_{ij}\leq c'_{ij}$ for all $ij,j$, then $\rho(C)\leq \rho(C')$.
Our main result shows that the assumption $C\leq C'$ can be a little loosen. 
Our theory is based on the following theorem, which is from \cite{chang}.

@@

    to the right of both terms of~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu,
    \end{equation}
 where the above equality follows by $Qu$ being eigenvector of $C'$ for $\lambda'$.
    Multiplying the nonnegative vector $v^T$ of $C$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu,
    \end{equation}
    where the above first equality follows by $v^TP$ being
    left eigenvector of $C$ for $\lambda$.
    Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies
        \begin{align*} &\lambda v^TPCQu=v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j\\ =&
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu=\lambda'v^TPC'Qu,
         \end{align*} so $\lambda=\lambda'$ by (\ref{pre2}).
\end{proof}

@@

We will apply Theorem~\ref{pre_thm} by using two particular square matrices $P$ and $Q$ to obtain our main result. 
We use $[n-1]$ as notation of the set of integers from $1$ to $n-1$, which is $\{1,2,\cdots,n-1\}$.
Throughout this thesis we fix $k\in [n-1]$, let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$. Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix}
1 &   &        &       &        &   & 0 \\
  & 1 &        &       &        &   &   \\
  &   & \ddots &       &        &   &   \\
  &   &        & 1     &        &   & 1 \\
  &   &        &       &\ddots  &   &  \\
  &   &        &       &        &1  & 0 \\
0 &   &        &       &        &0  & 1
\end{pmatrix},
\end{equation}
so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is
\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denotes the $(i,j)$-entry of $C'$.

@@

\begin{proof}
(i), (ii) follow immediate from the definition of $k$-rooted and
 $Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$.
\end{proof}

Below is our first result, in which the first condition implies the first $n-1$ columns of $C'$ major
 the same columns of $C$, and the sum of $k$-th and $n$-th columns of $C'$ also majors that of $C$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative $k$-rooted eigenvectors. And the forth condition is simpler
  but with the same meaning in Theorem~\ref{pre_thm}

@@

\begin{proof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}).
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem.
    Let $u = Q^{-1}v'$. Then $u$ is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and \ref{lem:rt_vec:en1} in
     Lemma~\ref{lem:rt_vec}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii}, \ref{thm_main:condition_iv} of this theorem since $P = I$ and
       $v'= Qu$. Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover,
        $\lambda = \lambda'$ if and only if (\ref{pre0}) holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a}, \ref{thm_main:equ_cond_b} and \ref{thm_main:equ_cond_c} of this theorem.
\end{proof}

@@

\begin{proof}
Since $$Q(I-E_{kn}=(I+E_{kn})(I-E_{kn})=I-E_{kn}+E_{kn}-E_{kn}E_{kn}=I,$$ we have
$Q^{-1}=I-E_{kn}$.
\end{proof}

@@

\begin{thm}\label{thm:conclusion}
    Let $C$ be an $n\times n$ nonnegative irreducible matrix and $C'$ be an $n\times n$ $k$-rooted matrix such that $C'Q$ majorizes $CQ$. 
 Then $\rho(C)\leq \rho(C')$. 
\end{thm}

@@

\draw[edge] (1) to[bend right] (3);
\draw[edge] (1) to[bend right] (4);
\draw[edge] (2) to (1);
\draw[edge] (2) to[bend right] (4);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend right] (1);
\draw[edge] (3) to[bend right] (4); 
\draw[edge] (4)[bend right] to (1); 
\draw[edge] (4)[bend right] to (2); 
\draw[edge] (4)[bend right] to (3); 
\end{tikzpicture}
\end{center}

@@

\draw[edge] (1) to[bend right] (3);
\draw[edge] (1) to[bend right] (4);
\draw[edge] (2) to (1);
\draw[edge] (2) to (3);
\draw[edge] (3) to[bend right] (1);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend right] (4);
\draw[edge] (4)[bend right] to (1);
\draw[edge] (4) to (2);
\draw[edge] (4)[bend right] to (3);
\end{tikzpicture}
\end{center}

Note that both $G$ and $G'$ have the same number of edges, 
neither $A$ majories $A'$ nor $A'$ majories $A$. Our main result still can do comparison between 
$\rho(A)$ and $\rho(A')$. 
We first observe that $A$ is irreducible and $A'$ is $k$-rooted for $k=3$. Moreover 
$A'Q=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}\geq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}=AQ$. 
Hence Hence $\rho(A)\leq\rho(A')$ by Theorem~\ref{thm:conclusion} with $C=A$ and $C'=A'$. 
Both of the values of $\rho(C)$ and $\rho(C')$ are close to $2.511547$ by  calculating on computer. \cite[sage]{sage}
\end{exam}

@@

All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented.\cite{src_files}

@@

@@

\bibitem{spec_rad}
A. E. Brouwer, W. H. Haemers, {\it Spectra of graphs}, Springer, 2012
\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
\bibitem{sage}
[Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
       The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial
\bibitem{src_files}
[github] K. H. Chen, C. W. Weng, source files to the essay, A simple method on comparison between spectral radii of two directed graphs,
     2021, https://reurl.cc/6yy32Z. % https://github.com/Urbaner3/NCTU_weng.

>>>>

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
 \usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman
\usepackage{graphicx}
\usepackage{comment}
%\usepackage{lipsum}
\usepackage{background}
\backgroundsetup{scale=1,angle=0,opacity=1,
    contents = {\includegraphics[scale = 1]{logo.jpg}}
}
%\usepackage{watermark}
%\watermark{\includegraphics[scale = 1]{logo.jpg}}

@@

%ij index orientation to your instinct, and do the definition

@@

The spectral radius is an important indicator to specify the relation
 of connected vertices in a graph, so it is meaningful to find a simple
 method to estimate the spectral radius. A simple and excellent
 executable method to estimate the spectral radius has some features.
 First, the bios is minimized, and second, there must be a way to prove
 it sensible. Enumerate these factors and prove it correctly would make this method reliable.

In \cite{chang}, Cheng and Weng give many bounds of the spectral radius of a nonnegative square matrix. And based on their theory and Perron-Frobenius theorem, we give another approach to obtain an upper bound of the spectral radius and apply it on the adjacency matrix of a directed graph.
All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented\cite{src_files}.

%張耀祖：有沒有比較何者較好？

@@

 A well-known application of Theorem~\ref{thm:Perron_Frobenius}
 show that if matrix  $C'$ {\it majors} $C$ (in notation $C\leq C'$),
 i.e. $c_{ij}\leq c'_{ij}$ for all $i,j$, then $\rho(C)\leq \rho(C')$.
Our main result shows that the assumption $C\leq C'$ can be a little loosen. 
Our theory is based on the following theorem, which is from \cite{chang}.

@@

    to the right of both terms of~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu,
    \end{equation}
 where the above equality follows by $Qu$ being eigenvector of $C'$ for $\lambda'$.
    Multiplying the nonnegative vector $v^T$ of $C$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu,
    \end{equation}
    where the above first equality follows by $v^TP$ being
    left eigenvector of $C$ for $\lambda$.
    Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies
        \begin{align*} &\lambda v^TPCQu=v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j\\ =&
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu=\lambda'v^TPC'Qu,
         \end{align*} so $\lambda=\lambda'$ by (\ref{pre2}).
         %This paragraph is hard to see the fact of proof. The formula is too brief.
\end{proof}

@@

We will apply Theorem~\ref{pre_thm} by using two particular square matrices $P$ and $Q$ to obtain our main result. 
We use $[n-1]$ as notation of the set of integers from $1$ to $n-1$, which is $\{1,2,\cdots,n-1\}$.
Throughout this thesis we fix $k\in [n-1]$. Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$. Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

%張：在Q的符號中看不出與k有關
\begin{equation} \label{Q_1}
Q=I+E_{kn}=\begin{pmatrix}
1 &   &        &       &        &   & 0 \\
  & 1 &        &       &        &   &   \\
  &   & \ddots &       &        &   &   \\
  &   &        & 1     &        &   & 1 \\
  &   &        &       &\ddots  &   &  \\
  &   &        &       &        &1  & 0 \\
0 &   &        &       &        &0  & 1
\end{pmatrix},
\end{equation}
so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is
\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denotes the $(i,j)$-entry of $C'$.

@@

\begin{proof}
(i), (ii) follow immediately from the definition of $k$-rooted and
 $Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$.
\end{proof}

Below is our first result, in which the first condition implies the first $n-1$ columns of $C'$ major
 the same columns of $C$, and the sum of $k$-th and $n$-th columns of $C'$ also majors that of $C$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative $k$-rooted eigenvectors. And the forth condition is simpler
  but with the same meaning in Theorem~\ref{pre_thm}.

@@

\begin{proof}
    The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}).
    The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem.
    Let $u = Q^{-1}v'$. Then $u$ is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and \ref{lem:rt_vec:en1} in
     Lemma~\ref{lem:rt_vec}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
      of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii}, \ref{thm_main:condition_iv} of this theorem since $P = I$ and
       $v'= Qu$. Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover,
        $\lambda = \lambda'$ if and only if (\ref{pre0}) holds, and this is equivalent to
         conditions \ref{thm_main:equ_cond_a}, \ref{thm_main:equ_cond_b}, and \ref{thm_main:equ_cond_c} of this theorem.
\end{proof}

@@

\begin{proof}
Since $$Q(I-E_{kn})=(I+E_{kn})(I-E_{kn})=I-E_{kn}+E_{kn}-E_{kn}E_{kn}=I,$$ we have
$Q^{-1}=I-E_{kn}$.
\end{proof}

@@

\begin{thm}\label{thm:conclusion}
    Let $C$ be an $n\times n$ nonnegative irreducible matrix and $C'$ be an $n\times n$ $k$-rooted matrix such that $C'Q$ majors $CQ$. 
 Then $\rho(C)\leq \rho(C')$. 
\end{thm}

@@

\draw[edge] (1) to (2);
\draw[edge] (1) to[bend left] (3);
\draw[edge] (1) to[bend left] (4);
\draw[edge] (2) to (3);
\draw[edge] (2) to[bend left] (4);
\draw[edge] (3) to[bend left] (1);
\draw[edge] (3) to[bend left] (4); 
\draw[edge] (4)[bend left] to (1); 
\draw[edge] (4)[bend left] to (2); 
\draw[edge] (4)[bend left] to (3); 
\end{tikzpicture}
\end{center}

@@

\draw[edge] (1) to (2);
\draw[edge] (1) to[bend left] (3);
\draw[edge] (1) to[bend left] (4);
\draw[edge] (2) to (3);
\draw[edge] (2) to (4);
\draw[edge] (3) to[bend left] (1);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend left] (4);
\draw[edge] (4)[bend left] to (1);
\draw[edge] (4)[bend left] to (3);
\end{tikzpicture}
\end{center}

Note that both $G$ and $G'$ have the same number of edges, 
neither $A$ majors $A'$ nor $A'$ majors $A$. Our main result still can do comparison between 
$\rho(A)$ and $\rho(A')$. 
We first observe that $A$ is irreducible and $A'$ is $k$-rooted for $k=3$. Moreover 
$A'Q=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}\geq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}=AQ$. 
Hence $\rho(A)\leq\rho(A')$ by Theorem~\ref{thm:conclusion} with $C=A$ and $C'=A'$. 
Both of the values of $\rho(C)$ and $\rho(C')$ are close to $2.511547$ by  calculating on computer \cite[sage]{sage}.
\end{exam}

@@

All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented \cite{src_files}.

@@

% 傅：tournaments graph, choosing the vertex with the maximum out-degree, does the theory implies the spectual radii are in an indecreasing sequence?  我：The irreducible in preserved.

@@

\bibitem{spec_rad}
A. E. Brouwer, W. H. Haemers, {\it Spectra of graphs}, Springer, 2012
\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
\bibitem{sage}
[Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
       The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial
\bibitem{src_files}
[github] K. H. Chen, source files to the essay, A simple method on comparison between spectral radii of two directed graphs,
     2021, https://reurl.cc/6yy32Z.  % https://github.com/Urbaner3/NCTU_weng.

     \newblock the link: $https://github.com/Urbaner3/NCTU\_weng$
